---
title: "Apply FungaLight to _S. cerevisiae_ treated with H2O2, compare with CFU"
author: Hanxi Tang, Bin He
date: "2025-03-05 (updated `r Sys.time()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
#require(ggrdiges)
```

```{r}
old <- theme_set(theme_minimal(base_size = 14) + panel_border(color = "gray20"))
```

# Background & Goal
Having optimized the FungaLight assay in _C. glabrata_ under hydrogen peroxide treatment and compared its result to CFU, we would like to test its applicability beyond this single species. To do so, Hanxi applied FungaLight to _S. cerevisiae_  and _C. albicans_, two other commonly used yeast species, also treated with hydrogen peroxide, and compared the results to CFU. Because the three species have different basal resistance to hydrogen peroxide, a different concentration series were used.

The experiments for _S. cerevisiae_ were conducted on 2024.02.25-28, and for _C. albicans_ on 20240318-23. Post-treatment cells were stained with PI or PI+SYTO9 (=FungaLight), and run through flow cytometry. The same sample was also plated for CFU.

This notebook will focus on the _S. cerevisiae_ data.

# Import data
> FCS files are stored in rdss/Users/Hanxi/flow cytometry under the above dates. They are copied to this repository under the same names.

_note_: samples are labeled by the H2O2 doses they were treated with. `_b` in the filename indicates the sample was stained with both PI and SYTO9, i.e., FungaLight, while `_p` indicates PI alone.


```{r}
# use relative path to make it easier for collaboration
data.path = "../input/20240225-28-h2o2-log-Sc-FGL/"
dat0 <- read.flowSet(path = data.path, pattern = "*.fcs",
                     transformation = FALSE,  # the original values are already linearized. 
                     emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                     column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
oriNames <- sampleNames(dat0)
tmp <- str_split(oriNames, pattern = "_", simplify = TRUE)[,c(1, 5, 6)]
colnames(tmp) <- c("Date", "Treatment", "Dye")
treatment.levels <- c("5C", "0", "0.6", "6", "1000")
sample <- data.frame(tmp) %>% 
  mutate(
    Dye = ifelse(Dye == "p.fcs", "PI", ifelse(Dye == "b.fcs", "Both", Dye)), 
    Treatment = fct_inorder(Treatment),
    name = paste(Date, 
                 ifelse(Treatment == "1000", "1M", paste0(Treatment, "mM")), 
                 Dye, sep = "_")
  )
rownames(sample) <- oriNames
dat <- dat0 # make a copy
pData(dat) <- sample
write_tsv(pData(dat), file = "../input/20240225-log-h2o2-sc-sample.tsv")
```


# Gatting strategies {.tabset}
## Load all gates from specification file
```{r}
source("../script/20250206-FGL-flow-gate-specification-Cg.R")
# below were the content initially used to populate the specification file above
## outlier gate
#outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
## single cell gate
#polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
#                    60, 105, 135,60), ncol = 2)
#colnames(polygon) <- c("FSC.H", "FSC.W")
#singlet.gate <- polygonGate(filterId = "singlet", .gate = polygon)
## live cell gate
#polygon <- matrix(c(0, 5*10^3, 5*10^3, 0, # BL1.H, green
#                    10^2.2, 10^2.2, 0, 0),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#live.gate <- polygonGate(filterId = "live", .gate = polygon)
## intermediate gate
#polygon <- matrix(c(10^3, 10^5.5, 10^5.5,   # BL1.H, green
#                    10^2.2, 10^2.2, 10^4.5),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#inter.gate <- polygonGate(filterId = "inter", .gate = polygon)
```

## Gating for singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.
>Gate for outlier 

```{r}
#outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat[1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat) # create a GatingSet
# rename the samples
sampleNames(gs) <- pData(gs)$name
# repair the Date column in pData
pData(gs)$Date = paste0("0", pData(gs)$Date)
# add the outlier gate
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

> Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat[[1]], outlier.gate)
#polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
#                    60, 105, 135,60), ncol = 2)
#colnames(polygon) <- c("FSC.H", "FSC.W")
#singlet.gate <- polygonGate(filterId = "singlet", .gate = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

## Gate for live cells
This is the live population with red fluorescence below 10^2.2

> **Main takeaway**
>
> 1. From 10mM to 1M treated samples, %live decreases and the change is much more obvious in the dual stain (PI first or simultaneous) than PI alone
>
> 2. The mock treated sample has a visible high red high green (not as high and prominent as the 100 mM treated) than 10 mM treated. It doesn't follow the trend above.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5.5), y = c(0,10^5.5)))
#polygon <- matrix(c(0, 5*10^3, 5*10^3, 0, # BL1.H, green
#                    10^2.2, 10^2.2, 0, 0),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#live.gate <- polygonGate(filterId = "live", .gate = polygon)
p.axis <- list(
  scale_x_logicle(breaks = 10^c(2,3,4,5)),
  scale_y_logicle(breaks = 10^c(2,3,4,5)),
  theme(
    strip.text = element_text(size = rel(1.1), face = 2)
  )
)
for(date in c("022524", "022624", "022824")){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(live.gate.sc) + 
    geom_stats(location = "gate", adjust = c(0.02, 2), digits = 1, ) + 
    labs(title = paste0("Date: ", date)) +
    facet_grid(Dye ~ factor(Treatment, levels = treatment.levels)) +# scPars +
    p.axis
  print(p)
}
```

Add this gate to the gatingSet

```{r}
gs_pop_add(gs, live.gate.sc, parent = "singlet", name = "live")
recompute(gs)
```
## Gate for the intermediate population with high SYTO9
This is the intermediate population with high Red and Green. We hypothesize that they represent oxidatively damaged cells with partially compromised plasma membranes, resulting in more SYTO9 accumulation but not a significant increase in PI.

```{r}
#polygon <- matrix(c(10^5, 1*10^3, 10^5,
#                    10^2.2, 10^2.2, 10^4), ncol = 2)
#polygon <- matrix(c(10^3, 10^5.5, 10^5.5,   # BL1.H, green
#                    10^2.2, 10^2.2, 10^4.5),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#inter.gate <- polygonGate(filterId = "inter", .gate = polygon)
for(date in c("022524", "022624", "022824")){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H),
              subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(inter.gate.sc) + 
    geom_stats(location = "data",  adjust = c(0.05, 5), digits = 3) + 
    labs(title = paste0("Date: ", date)) +
    facet_grid(Dye ~ factor(Treatment, levels = treatment.levels)) + scPars + 
    p.axis
  print(p)
}
```

```{r}
gs_pop_add(gs, inter.gate.sc, parent = "singlet", name = "intermediate")
recompute(gs)
```

## Representative FGL staining pattern
We have demonstrated our general gating strategy in Figure 4. Here, the goal is to show an example of the FGL staining pattern applied to other species. For _C. albicans_ in particular, there are some notable differences that required us to adjust the gate boundaries. We will skip PI alone staining for brevity and only show one replicate.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5), y = c(0,10^4.2)))
treatment_labels <- c(
  "0" = "0 mM", "0.6" = "0.6 mM", "6" = "6 mM", "1000" = "1M"
)
p <- ggcyto(gs[with(pData(gs), 
                    Date == "022824" & Dye == "Both" & Treatment != "5C")], 
            aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
  geom_hex(bins = 128) + 
  geom_gate(live.gate.sc, colour = "gray20") + 
  geom_gate(inter.gate.sc, colour = "steelblue3") + 
  #geom_stats(location = "gate", adjust = c(0.02, 2), digits = 1, fill = NA) + 
  facet_grid(. ~ factor(Treatment, levels = names(treatment_labels), 
                        labels = treatment_labels)) + 
  scPars + p.axis +
  theme(plot.title = element_blank(), axis.title = element_blank(),
        strip.text = element_text(size = rel(0.8), face = 2))
p

ggsave("../output/20250316-Sc-FGL-representative-plot.png", 
       height = 2, width = 5.5)
```

## Explore cell size difference in hydrogen peroxide treated cells
This is based on Jinye's suggestion:
> I am wondering is whether the cell size and granularity would affect the staining in different ways among species. Thus, I think it makes sense to plot the FL on FSC or SSC and perform a linear regression analysis for each species, if there is no significant difference, we can say the cell property of different species does not affect the staining in different ways. This is a simplified or adapted idea from PMID: 21952134.

Let's plot the distribution of FSC.H as a proxy for cell size
```{r}
mult_format <- function() {
     function(x) format(x/10000,digits = 2) 
}
```

```{r}
subset <- with(pData(gs), Treatment != "5C")
p1 <- ggcyto(gs[subset], aes(x = FSC.H), subset = "singlet") + 
  ggridges::geom_density_ridges(
    aes(y = factor(Treatment, levels = treatment.levels)),
        fill = "forestgreen", alpha = 0.8) + 
  scale_x_continuous(labels = mult_format(), name = "FSC.H x 10000") +
    facet_grid(Date ~ Dye) + labs(title = NULL)
p1
ggsave("../output/20250310-Sc-log-h2o2-cell-size-distribution-compare.png",
       width = 4, height = 4)
```

# Extract gated stats and MFI

% of events in each of the three gates
```{r}
# we are only interested in the populations after singlet gating
nodes <- c("singlet", "intermediate", "live")
# get the event count
gated_cnt <- gs_pop_get_stats(gs, nodes, type = "count")
# get the MFI for each parameter in each population of interest
gated_mfi <- gs_pop_get_stats(gs, nodes, type = pop.MFI) %>% 
  select(sample, pop, starts_with("FungaLight"))
```

Combine the data and meta data
```{r}
gated_stats <- full_join(gated_cnt, gated_mfi, by = c("sample", "pop")) %>% 
  right_join(pData(gs), by = c("sample" = "name")) %>% 
  relocate(Date, Treatment, Dye, .after = sample) %>% 
  relocate(sample, .after = last_col())
write_tsv(gated_stats, file = "../output/20250310-Sc-log-h2o2-gated-stats.tsv")

# we don't need the MFI for the latter analysis. add meta data to gated_cnt
gated_freq <- left_join(sample, gated_cnt, by = c("name" = "sample")) %>% 
  select(-name) %>% 
  group_by(Date, Treatment, Dye) %>% 
  mutate(perc = count / count[pop == "singlet"], .after = count)
```

```
#gated_stats <- gs_pop_get_stats(gs, type = "count") %>% 
#  as_tibble() %>% 
#  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
#  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count) %>% 
#  mutate(
#    #p_singlet = n_singlet / n_cells,
#    p_live = n_live / n_singlet,
#    p_intermediate = n_intermediate / n_singlet,
#    p_dead = 1 - p_live - p_intermediate
#  )
#write_tsv(gated_data, file = "../Input/20240124-fig-2-Dye-Order-gated-ht.tsv")
#print(gated_data)
```

```
final <- select(pData(dat), sample = name, everything()) %>% 
  left_join(select(gated_stats, -starts_with("n_")),  by = "sample") %>% 
  left_join(popMFI, by = "sample")

write_tsv(final, file = "../output/12.03-gated-output-20240214.tsv")
print(final %>% select(-sample))
```

# Compare with CFU
## Read CFU data 
These are averaged CFU provided by Hanxi. Here are the raw cfu data
```{r}
cfu_raw <- read_tsv("../input/20240225-0323-h2o2-Sc-Ca-CFU-raw.tsv",
                    col_types = "cccccciiii", comment = "#")
cfu_calc <- cfu_raw %>% 
  select(Date, Species, Treatment = H2O2_mM, Group, Dilution, Raw_count = Total) %>% 
  mutate(Date = ymd(Date) |> format("%m%d%y"),
         Treatment = fct_inseq(Treatment),
         CFU = Dilution * Raw_count) %>% 
  # group by Date to calculate the ratio of MO to MM
  group_by(Date) %>% 
  mutate(p_live = CFU / CFU[Group == "MM"]) %>% 
  # retain only Sc data
  dplyr::filter(Species == "Sc")

# export the data for paper
# write_tsv(cfu_calc, file = "../output/20250315-Sc-Ca-h2o2-cfu-estimate.tsv")
```

## Calculate survival from PI alone or FungaLight

For both PI alone and FungaLight, we will calculate the percent live events. For FungaLight specifically, we will also record the percent intermediate events and will investigate its utility
```{r}
dye_based <- gated_freq %>% 
  # record the total number of events as the count of singlets
  group_by(Date, Treatment, Dye) %>% 
  mutate(total_event_count = count[pop == "singlet"]) %>% 
  # now we don't need the singlet population and the count variable
  select(-count) %>% 
  dplyr::filter(pop != "singlet") %>%
  #ungroup() %>% 
  # we will pivot this data frame to a wider format, making 
  # two new variables out of the percentage variable
  pivot_wider(names_from = pop, values_from = perc, names_prefix = "p_") %>% 
  mutate(p_dead = 1 - p_live - p_intermediate) %>% 
  relocate(p_live, .after = total_event_count)

# export the data for paper
write_tsv(dye_based, file = "../output/20250314-Sc-h2o2-fgl-estimate.tsv")
```

```
PI <- final %>% 
  dplyr::filter(Dye == "PI") %>% 
  select(Date, Treatment, Method = Dye, Score = p_live)

FGL <- final %>% 
  dplyr::filter(Dye == "Both") %>% 
  mutate(
    wFGL = p_live * 2/log(BL1_live, 10)
  ) %>% 
  select(Date, Treatment, wFGL, FGL = p_live) %>% 
  pivot_longer(cols = c(wFGL, FGL), names_to = "Method", values_to = "Score")
  

all <- bind_rows(
  PI, FGL,
  mutate(cfu, Method = "CFU") %>% rename(Score = CFU)
) 
```

Combine the dye based results with cfu
```{r}
cfu_dye <- bind_rows(
  cfu_calc |> #dplyr::filter(Species == "Sc") |>
    mutate(Method = "CFU") |> 
    select(Date, Treatment, Method, p_live),
  select(dye_based, Date, Treatment, Method = Dye, starts_with("p"))) %>% 
  mutate(
    Method = factor(Method, levels = c("CFU", "PI", "Both"),
                    labels = c("CFU", "PI", "FungaLight")),
    Treatment = fct_inorder(Treatment)
  )
```

## Plotting the results
CFU survival
```{r}
p.cfu <- cfu_calc %>% 
  dplyr::filter(Species == "Sc") %>%
  ggplot(aes(x = Treatment, y = p_live)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray70", width = 0.8) +
  geom_point(aes(shape = as.character(Date)), 
             position = position_jitter(0.1),
             size = 1.5) + 
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  labs(title = "CFU", x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  theme(strip.text = element_text(size = rel(1.2), face = 2),
        strip.background = element_blank(),
        plot.title = element_text(hjust = 0.5))

p.cfu
```

FungaLight results
```{r}
p.dye <- dye_based %>% 
  dplyr::filter(Treatment != "5C", Dye == "Both") %>%
  ggplot(aes(x = fct_inseq(Treatment), y = p_live)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray70", width = 0.8) +
  geom_point(aes(shape = as.character(Date)), 
             position = position_jitter(0.1),
             size = 1.5) + 
  # overlay the p_intermediate points
  stat_summary(aes(y = p_intermediate), geom = "point", shape = 95,
            fun = "mean", size = 5, color = "steelblue3") +
  stat_summary(aes(y = p_intermediate, group = NA), geom = "line",
            fun = "mean", color = "steelblue") +
  geom_point(aes(y = p_intermediate, shape = as.character(Date)),
             position = position_jitter(0.1),
             color = "steelblue", size = 1.5, stroke = 1.5) + 
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  labs(title = "FungaLight", x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  theme(strip.text = element_text(size = rel(1.2), face = 2),
        strip.background = element_blank(),
        plot.title = element_text(hjust = 0.5))

p.dye
```

Alternative bar plots
```{r}
p.dye.side_by_side <- dye_based %>% 
  dplyr::filter(Treatment != "5C", Dye == "Both") %>%
  select(Date, Treatment, p_live, p_intermediate) %>% 
  pivot_longer(cols = c(p_live, p_intermediate), 
               names_to = "par", values_to = "perc") %>% 
  mutate(par = factor(par, levels = c("p_live", "p_intermediate"),
                      labels = c("% live", "% intermediate"))) %>% 
  ggplot(aes(x = fct_inseq(Treatment), y = perc, group = par)) +
  geom_bar(aes(fill = par), position = position_dodge(0.9),
           stat = "summary", fun = "mean", width = 0.8) +
  geom_point(aes(shape = as.character(Date), color = par), 
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.9),
             size = 1.2) + 
  scale_fill_manual(NULL, values = c("gray70", "lightblue2")) +
  scale_color_manual(NULL, values = c("black", "steelblue")) +
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  labs(title = "FungaLight", x = bquote(H[2]*O[2]~(mM)), y = NULL) +
  theme(strip.text = element_text(size = rel(1.2), face = 2),
        strip.background = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "inside",
        legend.position.inside = c(0.8, 0.9))

p.dye.side_by_side
```

Assemble the plot
```{r}
plot_grid(
  p.cfu + labs(x = NULL, y = NULL) +
    theme(plot.title = element_text(size = rel(0.7),
                                    margin = margin(b = 2)),
          axis.text = element_text(size = rel(0.7)),
          axis.text.x = element_blank()),
  NULL,
  p.dye.side_by_side + labs(x = NULL, y = NULL) +
    theme(#axis.text.y = element_blank(),
          plot.title = element_text(size = rel(0.7),
                                    margin = margin(b = 1)),
          axis.text = element_text(size = rel(0.7)),
          legend.position = "none"),
  nrow = 3, rel_heights = c(1, 0.1, 1.12)
)
ggsave("../output/20250321-Sc-log-h2o2-CFU-FGL-compare.png",
       width = 2.5, height = 3)
```

Compare the three methods
```{r}
#h2o2.levels <- c("0" = "0", "10" = "10", "100" = "100", "1000" = "1K")
p <- cfu_dye %>% 
  dplyr::filter(Treatment != "5C") %>% 
  ggplot(aes(x = Treatment, y = p_live, group = Method)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray70", width = 0.7) +
  geom_point(aes(shape = Date), position = position_jitter(0.2),
             size = 1.5) + 
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  scale_x_discrete() +
  #coord_flip() + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  facet_wrap(~Method, nrow = 1) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = rel(0.9), face = 1),
        axis.text.x = element_text(size = rel(0.9), angle = 90, vjust = .5),
        axis.text.y = element_text(size = rel(0.9)),
        axis.title = element_text(size = rel(0.9)),
        panel.spacing = unit(1, "lines"))
p
ggsave("../output/20250315-Sc-perc-live-compare-three-methods.png",
       width = 4, height = 2.5)
```

## Statistical tests
```{r}
tmp <- aov(p_live ~ Treatment, data = cfu_calc)
TukeyHSD(tmp)
```

For p_live with FungaLight
```{r}
tmp <- dye_based %>% 
  dplyr::filter(Treatment != "5C", Dye == "Both") %>%
  aov(p_live ~ fct_inseq(Treatment), data = .)
TukeyHSD(tmp)
```

For p_intermediate with FungaLight
```{r}
tmp <- dye_based %>% 
  dplyr::filter(Treatment != "5C", Dye == "Both") %>%
  aov(p_intermediate ~ fct_inseq(Treatment), data = .)
TukeyHSD(tmp)
```

Below we use ANOVA and posthoc test to test for significant differences between the conditions
```{r}
#tmp <- dplyr::filter(cfu_dye, Treatment != "5C", Method == "FungaLight") %>% 
#  mutate(Treatment = fct_inseq(Treatment))
#lm <- aov(p_live ~ Treatment, data = tmp)
#TukeyHSD(lm)

cfu_dye %>% ungroup() %>% 
  dplyr::filter(Treatment != "5C") %>% 
  mutate(Treatment = fct_inseq(Treatment)) %>% 
  select(Treatment, Method, p_live) %>% 
  nest(data = c(Treatment, p_live)) %>% 
  mutate(
    fit = map(data, \(df) aov(p_live ~ Treatment, data = df)),
    posthoc = map(fit, TukeyHSD),
    tidied = map(posthoc, broom::tidy)
  ) %>% 
  unnest(tidied) %>% 
  select(Method, contrast, diff = estimate, lwr = conf.low, upr = conf.high,
         Padj = adj.p.value)
```

---

Below are hanxi's plotting code
<!---
```
all %>% 
  dplyr::filter(!Treatment %in% c("mock", "5C"), Method != "wFGL") %>% 
  mutate(Method = fct_relevel(Method, "CFU", "FGL", "PI")) %>% 
  ggplot(aes(x = Treatment, y = Score)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray") +
  #geom_point(size = 2, position = position_jitter(0.05)) + 
  stat_summary(fun.data = "mean_sdl", fun.args = list(mult = 1), 
               geom = "errorbar",  width = 0.1) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  facet_wrap(~Method, nrow = 1) +
  theme_cowplot() +
  theme(strip.text = element_text(size = rel(1.2), face = 2),
        strip.background = element_blank())
ggsave("../output/20240323-for-R35-CFU-compared-with-FGL-and-PI.png", width = 5, height = 3)
```

```
combined_data <- read_tsv("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/02-FungaLight-Distinguishing-Power/Input/20240219-PI-Both-Distinguishing-Power-gated-CFU-ht.tsv")
print(combined_data)
combined_data %>% 
  dplyr::filter(Dye %in% c("Both")) %>% 
  dplyr::filter(Treatment %in% c("0", "10", "100")) %>% 
  ggplot(aes(x = FGS, y = CFU_Survival)) +
  geom_point(aes(color = Treatment))

combined_data %>% 
  dplyr::filter(Dye %in% c("Both")) %>% 
  ggplot(aes(x = Treatment, y = FGS)) +
  geom_point(aes(color = Date))

combined_data %>% 
  dplyr::filter(Dye %in% c("Both")) %>% 
  ggplot(aes(x = Treatment, y = percent_inter)) +
  geom_point(aes(color = Date))
#ggsave("../output/20240104-PI-Dilutions-Gated_RP-Plot.png", width = 7, height = 5)
```


```
combined_data$new_FGS <- 1*combined_data$percent_Live - 0.6*combined_data$percent_inter - combined_data$percent_Dead
pf1g <- combined_data %>%
  pivot_longer(cols = c(CFU_Survival, new_FGS),
               names_to = "Assay",
               values_to = "Survival") %>% 
  dplyr::filter(!(Dye %in% c("PI"))) %>% 
  dplyr::filter(!(Treatment %in% c("5C")))  %>%
     mutate(Assay = factor(Assay, levels = c("CFU_Survival", "new_FGS"), labels = c("CFU", "FGL"))) %>% 
  #mutate(Dilution = factor(Dilution, levels = c("250 x", "500 x", "1000 x", "2000 x"))) %>%
  ggplot(aes(x = Assay, y = Survival)) +
  facet_wrap(~ Treatment, nrow = 1, labeller = as_labeller(c("0" = "0 mM", "10" = "10 mM", "100" = "100 mM", "1000" = "1000 mM"))) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red2",
               size = 0.8, position = position_nudge(x = 0.2)) +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.01)) +
  scale_y_continuous() +
  theme_minimal()

pf1g +
  labs(x = "Assay Type", y = "Survival Quantification") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 20, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 20, face = "bold", color = "black"),
        axis.title.x = element_text(size = 20, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240301-Cg-Log-Gated_.png", width = 7, height = 5)
```




```
combined_data$percent_PIDead <- 1 - (combined_data$n_PIDead / combined_data$n_singlet)
p <- combined_data %>%
  pivot_longer(cols = c(CFU_Survival, percent_PIDead),
               names_to = "PIAssay",
               values_to = "PISurvival") %>% 
  dplyr::filter(!(Dye %in% c("Both"))) %>% 
  dplyr::filter(!(Treatment %in% c("5C")))  %>%
     mutate(PIAssay = factor(PIAssay, levels = c("CFU_Survival", "percent_PIDead"), labels = c("CFU", "PI"))) %>% 
  #mutate(Dilution = factor(Dilution, levels = c("250 x", "500 x", "1000 x", "2000 x"))) %>%
  ggplot(aes(x = PIAssay, y = PISurvival)) +
  facet_wrap(~ Treatment, nrow = 1, labeller = as_labeller(c("0" = "0 mM", "10" = "10 mM", "100" = "100 mM", "1000" = "1000 mM"))) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red2",
               size = 0.8, position = position_nudge(x = 0.2)) +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.01)) +
  scale_y_continuous() +
  theme_minimal()

p +
  labs(x = "Assay Type", y = "Survival Quantification") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 20, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 20, face = "bold", color = "black"),
        axis.title.x = element_text(size = 20, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240301-Cg-Log-Gated-PI.png", width = 7, height = 5)
```

```
pf1g <- combined_data %>%
  dplyr::filter(!(Dye %in% c("Both"))) %>% 
  dplyr::filter(!(Treatment %in% c("1000", "5C")))  %>%
  #mutate(Dilution = factor(Dilution, levels = c("250 x", "500 x", "1000 x", "2000 x"))) %>%
  ggplot(aes(x = Treatment, y = CFU_Survival)) +
  facet_wrap(~ Dye, scale = "free_y") +
  geom_boxplot() +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
pf1g +
  labs(x = "Hydrogen Peroxide Treatment Concentration", y = "CFU Survival") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold", color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 16, face = "bold")) 
#ggsave("../output/20240104-PI-Dilutions-Gated_RP-Plot.png", width = 7, height = 5)
```
--->
