---
title: "PI vs PI+SYTO9 Log Correlation"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for post treatment phase Ca stained with PI/PI+SYTO9, and run through flow cytometry (details in ELN), and plated for CFU. The goal is to establish FungaLight's distinguishing power and correlate with plating based method.

```{r setup, message=FALSE}
library(tidyverse)
library(flowCore)
library(flowClust)
library(openCyto)
library(ggcyto)
library(cowplot)
require(ggridges)
library(PTXQC)
library(ggplot2)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "./input/20240318-0323_Ca_Log_FGL"
dat.f1g <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
oriNames <- sampleNames(dat.f1g)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 7, 8)]
colnames(tmp) <- c("Date", "Treatment", "Dye") 
sample <- data.frame(tmp) %>% 
  mutate(
    Dye = ifelse(Dye == "p.fcs", "PI", ifelse(Dye == "b.fcs", "Both", Dye)), 
    Treatment = factor(Treatment, levels = c(0, 0.6, 6, 1000)))
rownames(sample) <- oriNames
pData(dat.f1g) <- sample
print(pData(dat.f1g))
#write_tsv(pData(dat.f1g), file = "/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/02-FungaLight-Distinguishing-Power/Input/20240325-Ca-PI-Both-Log-Correlation-raw-data-ht.tsv")
```

# Correting green (BL1) and red (BL3) fluorescence by normalization to cell size, will move it after prepossing the data (gated out outliers and spurious events)(Creator: JY Liang)
## There is no BL1.A or BL3.A data, only BL1.H and BL3.H. I cannot use the formula: Normalization GFP by cell size: NormFL1.A = FL1.A * sqrt(1/(FSC.A*SSC.A))
## Instead, I can plot the BL1.H or BL3.H against the FSC and SSC (should choose area over height), then generate a regression model of BL1.H (or BL3.h) on SSC and SSC for each sample in Ca and the Sc (in other datasets), if there is no significant


# before gating out outliers and adjusted for singlets

```{r}
#Extract the first sample (flowFrame) matrix from the flowSet dat.f1g:
sample_data <- dat.f1g[[1]]
event_data <- as.data.frame(exprs(sample_data))


plot_scatter <- function(data, x_var, y_var, color, x_label, y_label, title) {
  ggplot(data, aes_string(x = x_var, y = y_var)) +
    geom_point(color = color, alpha = 0.5) +
    ggtitle(title) +
    xlab(x_label) +
    ylab(y_label) +
    theme_minimal()
}

plot_scatter(event_data, "FSC.H", "BL1.H", "darkgreen", "FSC.H (Cell Size)", "BL1.H (GFP Intensity)", "FSC.H vs BL1.H")

plot_scatter(event_data, "SSC.H", "BL1.H", "lightgreen", "SSC.H (Cell Granularity)", "BL1.H (GFP Intensity)", "SSC.H vs BL1.H")

plot_scatter(event_data, "FSC.H", "BL3.H", "darkred", "FSC.H (Cell Size)", "BL3.H (GFP Intensity)", "FSC.H vs BL3.H")

plot_scatter(event_data, "SSC.H", "BL3.H", "red2", "SSC.H (Cell Granularity)", "BL3.H (GFP Intensity)", "SSC.H vs BL3.H")



```




# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f1g [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f1g) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f1g[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```

Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```
> Visualize the distribution of cell size & the relationship between fluorescence and cell size parameters (Creator: JY)
Visualize cell size parameters with Fluorescence

```{r}
#removes outlier
filtered_outlier <- Subset(dat.f1g[[1]],outlier.gate)

#filter for singlet
filtered_singlets <- Subset(filtered_outlier, singlet.gate)

#Convert to dataframe
df1 <- as.data.frame(exprs(filtered_singlets)) %>% mutate(Treatment = "PI", date_info = "031824")

# visualize the gated distribution
plot_scatter(df1, "FSC.H", "BL1.H", "darkgreen", "FSC.H (Cell Size)", "BL1.H (GFP Intensity)", "FSC.H vs BL1.H")

plot_scatter(df1, "SSC.H", "BL1.H", "lightgreen", "SSC.H (Cell Granularity)", "BL1.H (GFP Intensity)", "SSC.H vs BL1.H")

plot_scatter(df1, "FSC.H", "BL3.H", "darkred", "FSC.H (Cell Size)", "BL3.H (GFP Intensity)", "FSC.H vs BL3.H")

plot_scatter(df1, "SSC.H", "BL3.H", "red2", "SSC.H (Cell Granularity)", "BL3.H (GFP Intensity)", "SSC.H vs BL3.H")
```
create density plot
```{r}
ggplot(df1,aes(x = FSC.H)) +
  geom_density(fill = "darkgreen", alpha = 0.5) +
  scale_x_continuous(name = "FSC.H x10,000",
                     breaks = seq(0,1000000, by = 250000), # set custom tick marks
                     labels = seq(0,100, by = 25)) + # convert original values to a scale of 0 - 100
  labs(y = "Treatment_031824_0.6_PI") +
  theme_minimal()

```


visualize the difference between _b and _p, what is there difference? b represents both dye, P represents PI only
```{r}
test_data <- dat.f1g[[2]]
df_test <- as.data.frame(exprs(test_data))



```


pull together density plot
```{r}
# Create a GatingSet from the flowSet
gs <- GatingSet(dat.f1g)

# Define the outlier gate (FSC.H and SSC.H gating)
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))

# Apply outlier gate to the GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)

# Define the singlet gate (FSC.H and FSC.W gating)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135, 60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)

# Apply singlet gate to the GatingSet
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)

# Function to extract metadata from sample name
extract_metadata <- function(filename) {
  date <- substr(filename, 1, 6)  # Extract first 6 characters as date
  treatment_code <- str_sub(filename, -5, -5)  # Extract the character before ".fcs"
  # Extract the H2O2 concentration (number before 'b' or 'p' and '.fcs')
  h2o2_match <- str_extract(filename, "(\\d*\\.?\\d+)(?=_b\\.fcs|_p\\.fcs)") 
  
  # Map treatment codes to labels
  treatment <- ifelse(treatment_code == "b", "Both",
                      ifelse(treatment_code == "p", "PI", "Unknown"))
  
  return(data.frame(Date = date, Treatment = treatment, H2O2_conc = as.numeric(h2o2_match)))
}


# Extract gated data for all samples
gated_data_list <- list()

for (i in seq_along(dat.f1g)) {
  sample_name <- sampleNames(dat.f1g)[i]  # Get filename
  metadata <- extract_metadata(sample_name)  # Extract metadata (date, treatment, H2O2)

  # Subset data using the gates
  filtered_outlier <- Subset(dat.f1g[[i]], outlier.gate)
  filtered_singlets <- Subset(filtered_outlier, singlet.gate)

  # Convert to dataframe and add metadata
  df <- as.data.frame(exprs(filtered_singlets)) %>%
    mutate(Treatment = metadata$Treatment, 
           Date = metadata$Date, 
           H2O2_conc = metadata$H2O2_conc)

  gated_data_list[[sample_name]] <- df  # Store in list
}

# Combine all gated data into a single dataframe
gated_data <- bind_rows(gated_data_list, .id = "Sample")

# Print first few rows of the final dataframe
head(gated_data)

```




```{r}
library(ggplot2)
library(ggridges)
library(dplyr)

# Scale FSC.H values to match x-axis in the example (FSC.H / 10,000)
gated_data <- gated_data %>%
  mutate(
    FSC.H_scaled = FSC.H / 10000,
    H2O2_conc = as.numeric(as.character(H2O2_conc))  # Ensure numeric
  )

# Create the density plot
#plot<- 
ggplot(gated_data, aes(x = FSC.H_scaled, y = factor(H2O2_conc), fill = ..density..)) +
  geom_density_ridges(scale = 2, alpha = 0.8, color = "black", fill = "green4") +
  facet_grid(rows = vars(Date), cols = vars(Treatment), scales = "free_y", space = "free_y") +
  scale_x_continuous(limits = c(0, 100), breaks = seq(25, 100, by = 25)) +
  theme_minimal() +
  labs(x = "FSC.H x 10,000", y = "Treatment") +
  theme(
    strip.text = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold", color = "black"),
    axis.text.x = element_text(size = 12, face = "bold", color = "black"),
    panel.grid.major = element_line(color = "gray"),
    panel.grid.minor = element_blank()
  )

#save
#ggsave("output/Ca_PIvsPI_FGL.png", plot = plot, width = 10, height = 8, dpi = 300)

```



## Test Normalization
Since the 0.6mM H2O2 treatment for C. albicans leads to a right-shift cell size distribution (indicated by FSC.H).I want to calculate the if the cell size normalized nBL1.H and nBL3.H will have different distribution than the BL1.H and BL3.H.

Goal: examine if the cell size significantly affect the fluorescence across samples in different H2O2 treatment.

```{r}
# calculate the median of FSC.H from samples grouped in 0mM, 0.6mM, 6mM and 1000mM
# gated_data is the singlets gated raw data from dat.f1g

mfsc <- gated_data %>% 
        group_by(H2O2_conc, Treatment) %>% 
        summarise(median_FSC.H = median(FSC.H, na.rm = TRUE),
                  count = n())
exponent <- 1.2
```
```{r}
# calculate ratio
norm_data <- gated_data %>% 
             left_join(mfsc, by = c("H2O2_conc", "Treatment")) %>% 
             mutate(nRFP = BL3.H/(FSC.H/median_FSC.H)^(exponent),
                    nGFP = BL1.H/(FSC.H/median_FSC.H)^(exponent))
```

# plot the density plot of BL1.H vs BL3.H and nGFP vs nRFP
Conclusion: estimated cell normalization on BL1.H and BL3.H channel does not affect dead cell population conclusion, based on gating on red positive (dead; for Both).

I can define a function that specify the channel with or w/o normaliztion, and specify the polygon for gating the cells.

# before normalization
```{r}
library(ggplot2)
library(dplyr)
library(sp)

# Define polygon with BL1.H and BL3.H
polygon <- matrix(c(10^3, 10^2.5, 10^3, 10^5,
                    10^4.5, 10^2.2, 10^2.2, 10^4), 
                  ncol = 2)  # Default fills column-wise
colnames(polygon) <- c("BL1.H", "BL3.H")

# Check if each point is inside the polygon
norm_data <- norm_data %>%
  mutate(inside_polygon = point.in.polygon(BL1.H, BL3.H, 
                                           polygon[, 1], polygon[, 2]) > 0)

# Summarize percentage inside
summary_counts <- norm_data %>%
  group_by(H2O2_conc, Treatment) %>%
  summarise(
    total_points = n(),
    inside_count = sum(inside_polygon, na.rm = TRUE),
    percentage_inside = round((inside_count / total_points) * 100, 3),  # Keep naming
    .groups = "drop"
  )

ggplot(norm_data, aes(x = BL1.H, y = BL3.H)) +
  geom_hex(bins = 80, aes(fill = ..count..)) +  # Hex binning
  geom_polygon(data = as.data.frame(polygon), aes(x = BL1.H, y = BL3.H), 
               fill = NA, color = "red", size = 1.2) +  # Polygon overlay
  geom_text(data = summary_counts, 
            aes(x = 10^3.5, y = 10^4.5,  # Adjusted placement
                label = paste0(percentage_inside, "%")), 
            color = "black", size = 6, fontface = "bold") +  # Add percentage text
  facet_grid(rows = vars(H2O2_conc), cols = vars(Treatment)) +
  theme_minimal() +
  labs(x = "BL1.H", y = "BL3.H") +
  scale_x_logicle() +  
  scale_y_logicle()



```

# nGFP vs nRFP
```{r}

# for nGFP vs nRFP
# Convert polygon matrix to a data frame
polygon_df <- as.data.frame(matrix(c(10^3, 10^2.5, 10^3, 10^5,
                                     10^4.5, 10^2.2, 10^2.2, 10^4), 
                                   ncol = 2))
colnames(polygon_df) <- c("nGFP", "nRFP")

# Check if each point is inside the polygon
norm_data <- norm_data %>%
  mutate(inside_polygon = point.in.polygon(nGFP, nRFP, 
                                           polygon_df$nGFP, polygon_df$nRFP) > 0)

# Summarize percentage inside
summary_counts <- norm_data %>%
  group_by(H2O2_conc, Treatment) %>%
  summarise(
    total_points = n(),
    inside_count = sum(inside_polygon, na.rm = TRUE),
    percentage_inside = round((inside_count / total_points) * 100, 3),  # Round for readability
    .groups = "drop"
  )

ggplot(norm_data, aes(x = nGFP, y = nRFP)) +
  geom_hex(bins = 80, aes(fill = ..count..)) +  # Hex binning
  geom_polygon(data = polygon_df, aes(x = nGFP, y = nRFP), 
               fill = NA, color = "red", size = 1.2) +  # Overlay polygon
  geom_text(data = summary_counts, 
            aes(x = 10^3.5, y = 10^4.5,  # Adjust position for visibility
                label = paste0(percentage_inside, "%")), 
            color = "black", size = 6, fontface = "bold") +  # Overlay text
  facet_grid(rows = vars(H2O2_conc), cols = vars(Treatment)) +
  theme_minimal() +
  labs(x = "nGFP", y = "nRFP") +
  scale_x_logicle() +  
  scale_y_logicle()

```
# Define a function that can specify the polygon, treatment, with or w/o normalization
```{r}
# get polygon on ggplot:
get_polygon <- function(stain, dat, live_dead) {
  # Initialize polygon_df as a data frame
  polygon_df <- data.frame(BL1.H = numeric(), BL3.H = numeric())
  
  # specify red population:
  if (live_dead == "red_neg") {
    # red negative population:
  polygon_df <- as.data.frame(matrix(c(0, 10^5, 10^5, 0, 10^2.5, 10^2.5, 0, 0), ncol = 2))
  colnames(polygon_df) <- c("BL1.H", "BL3.H")
  }
  else if (live_dead == "red_int") {
  # red intermediate population:
  polygon_df <- as.data.frame(matrix(c(10^5, 10^3,10^5,
                    10^2.2, 10^2.2, 10^4), ncol = 2))
  colnames(polygon_df) <- c("BL1.H", "BL3.H")
  }
  # red positive polulation:
  else {
  if (stain == "b") {
      polygon_df <- as.data.frame(matrix(c(10^3, 10^2.5, 10^3, 10^5,
                                     10^4.5, 10^2.2, 10^2.2, 10^4), 
                                   ncol = 2))
      colnames(polygon_df) <- c("BL1.H", "BL3.H")
  } 
  else if (stain == "p"){
    # red positive dead for PI only
    polygon_df <- as.data.frame(matrix(c(0, 10^3.6, 10^3.6, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2))
    colnames(polygon_df) <- c("BL1.H", "BL3.H")
  } 
  else {print("b or p for stain")}}
  
  return(polygon_df) 
}

# calculate percentage of cells within the polygon
get_polygon_pct <- function(stain, dat, live_dead) {
  # Initialize polygon_df as a data frame
  polygon_df <- data.frame(BL1.H = numeric(), BL3.H = numeric())
  
  # specify red population:
  if (live_dead == "red_neg") {
    # red negative population:
  polygon_df <- as.data.frame(matrix(c(0, 10^5, 10^5, 0, 10^2.5, 10^2.5, 0, 0), ncol = 2))
  colnames(polygon_df) <- c("BL1.H", "BL3.H")
  }
  else if (live_dead == "red_int") {
  # red intermediate population:
  polygon_df <- as.data.frame(matrix(c(10^5, 10^3,10^5,
                    10^2.2, 10^2.2, 10^4), ncol = 2))
  colnames(polygon_df) <- c("BL1.H", "BL3.H")
  }
  # red positive polulation:
  else {
  if (stain == "b") {
      polygon_df <- as.data.frame(matrix(c(10^3, 10^2.5, 10^3, 10^5,
                                     10^4.5, 10^2.2, 10^2.2, 10^4), 
                                   ncol = 2))
      colnames(polygon_df) <- c("BL1.H", "BL3.H")
  } 
  else if (stain == "p"){
    # red positive dead for PI only
    polygon_df <- as.data.frame(matrix(c(0, 10^3.6, 10^3.6, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2))
    colnames(polygon_df) <- c("BL1.H", "BL3.H")
  } 
  else {print("b or p for stain")}}
  
    # Check if each point is inside the polygon
  norm_data <- dat %>% mutate(inside_polygon = point.in.polygon(BL1.H, BL3.H, polygon_df$BL1.H, polygon_df$BL3.H) > 0)

   # Summarize percentage inside
  summary_counts <- norm_data %>%
                    group_by(H2O2_conc, Treatment) %>% summarise(
                    total_points = n(), inside_count = sum(inside_polygon, na.rm = TRUE),
                    percentage_inside = round((inside_count / total_points) * 100, 3),  # Round for readability
                    .groups = "drop")
  
  return(summary_counts)
}

test <- get_polygon_pct("b", norm_data, "red_neg")
test_polygon <- get_polygon("b", norm_data, "red_neg")

# need to specifically define a function only draw polygon; then define a function that get the polygon summary
```

```{r}
ggplot(norm_data, aes(x = BL1.H, y = BL3.H)) +
  geom_hex(bins = 80, aes(fill = ..count..)) +  # Hex binning
  geom_polygon(data = test_polygon, aes(x = BL1.H, y = BL3.H), 
               fill = NA, color = "red", size = 1.2) +  # Overlay polygon
  geom_text(data = test, 
            aes(x = 10^3.5, y = 10^4.5,  # Adjust position for visibility
                label = paste0(percentage_inside, "%")), 
            color = "black", size = 6, fontface = "bold") +  # Overlay text
  facet_grid(rows = vars(H2O2_conc), cols = vars(Treatment)) +
  theme_minimal() +
  labs(x = "BL1.H", y = "BL3.H") +
  scale_x_logicle() +  
  scale_y_logicle()

```



### DO NOT USE THEM
> Gate for Red Negative Population
This is the live population with red fluorescence below 10^2.2

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set (limits = list(x = c(0, 1e6), y = c(0, 1e6)))
ex <- Subset(dat.f1g[[1]], singlet.gate)
polygon <- matrix(c(0, 10^5, 10^5, 0,
                    10^2.5, 10^2.5, 0, 0), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
Nlive.gate <- polygonGate(filterId = "Nlive", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + scPars + scale_x_logicle() + scale_y_logicle()
```


Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, Nlive.gate, parent = "singlet", name = "Nlive")
recompute(gs)
```

> Gate for Red Positive Intermediates
This is the intermediate population on the right side

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(10^5, 10^3,10^5,
                    10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
inter.gate <- polygonGate(filterId = "inter", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(inter.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_grid (Treatment ~ Dye)
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, inter.gate, parent = "singlet", name = "inter")
recompute(gs)
```

> Gate for Red Positive Dead
This is the dead population on the right side

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(10^3, 10^2.5,10^3, 10^5,
                    10^4.5, 10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_grid (Treatment ~ Dye)
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```

> Gate for Dead Population - PI only
RP population is defined by having red fluorescence value above 10^2.15.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g[[1]], singlet.gate)
polygon <- matrix(c(0, 10^3.6, 10^3.6, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
PIdead.gate <- polygonGate(filterId = "PIdead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(PIdead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, PIdead.gate, parent = "singlet", name = "PIDead")
recompute(gs)
```

>Exported gated event counts

```{r}
gated_stats_f1g <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f1g$percent_Live <- gated_stats_f1g$n_Nlive / gated_stats_f1g$n_singlet
gated_data <- cbind(pData(dat.f1g), gated_stats_f1g) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
write_tsv(gated_data, file = "./input/20240704-Ca-PI-Both-Distinguishing-Power-gated-ht.tsv")
print(gated_data)
```
# Extract MFI of intermediate population green channel
>MFI are extracted and wrote into tsv. The tsv is stored in the input folder.

```{r}
interpop <- Subset(dat.f1g, inter.gate)
tmp <- fsApply(interpop, each_col, median)
combined_data <- cbind(gated_data, tmp) %>%
    as.tibble() %>% relocate(name, .after = last_col())

#write_tsv(combined_data, file = "./input/20240124_four_variable_matrix.tsv")
print(combined_data)
```


# Output Graph Production
## Before cell size normalization
>Several output graphs are produced, and are stored in the Output folder.

```{r}
sub <- dat.f1g[pData(dat.f1g)$Dye == "Both" & pData(dat.f1g)$Date == "032324"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Dye ~ Treatment, 
             labeller = as_labeller(c("0" = "0 mM", "0.6" = "0.6 mM", "6" = "6 mM", "1000" = "1000 mM",
                                       "PI" = "PI only", "SYTO9_first" = "SYTO9+PI Costain", "PI_first" = "PI first"))) + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 9, face = "bold")) 
#ggsave("./output/20240117-dye-Order-Flow-Plot.png", width = 7, height = 5)
```
```{r}
combined_data <- read_tsv("./input/20240325-Ca-PI-Both-Distinguishing-Power-gated-CFU.tsv") %>% 
mutate(Treatment = factor(Treatment, levels = c(0, "5C", 0.6, 6, 1000)))

p1 <- combined_data %>% 
  dplyr::filter(Dye == "PI") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Red Negative") +
  ggtitle("PI only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 16) 
  

p2 <- combined_data %>% 
  dplyr::filter(Dye == "Both") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = CFU_Survival))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Viable") +
  ggtitle("CFU") +  
  theme_cowplot(font_size = 16) 
  theme(plot.title = element_text(hjust = 15))

p3 <- combined_data %>% 
  dplyr::filter(Dye == "Both") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.1)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live Population") +
  ggtitle("FungaLight") +  
  theme(plot.title = element_text(hjust = 15)) +
  theme_cowplot(font_size = 16) 

combined_plot <- plot_grid(p1, p2, p3, nrow = 1, ncol = 3, common.legend = TRUE, align = 'hv')

print(combined_plot)
#ggsave("../output/20240427-Ca-Log-Gated.png")
```


```{r}
combined_data1 <- read_tsv("./input/20240619-Ca-PI-Both-Distinguishing-Power-gated-CFU.tsv")
print(combined_data1)
```


```{r}
p1 <- combined_data1 %>% 
  dplyr::mutate(Treatment = as.character(Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = c(0, 0.6, 6, 1000))) %>%
  dplyr::filter(!(Treatment %in% c("0"))) %>%
  ggplot(aes(x = Treatment, y = Survival)) + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  theme_cowplot(font_size = 20) +
  facet_wrap(~ Method)

p1

ggsave("./output/20240618-Ca-Log-Gated.png", width = 7, height = 4)
```

```{r}
combined_data1 <- read_tsv("C:/Users/tangh/Downloads/20240325-Ca-PI-Both-Distinguishing-Power-gated-CFU.tsv") # where it is
print(combined_data1)
```

```{r}
p1 <- combined_data1 %>% 
  dplyr::mutate(Treatment = as.character(Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = c(0, 0.6, 6, 1000))) %>%
  dplyr::filter(!(Treatment %in% c("0"))) %>%
  ggplot(aes(x = Treatment, y = Survival)) + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  theme_cowplot(font_size = 20) +
  facet_wrap(~ Method)

p1

ggsave("../output/20240704-Ca-Log-Gated.png", width = 7, height = 4)
```

Statistical Test

```{r}
# Filter the data
filtered_data <- combined_data1 %>% dplyr::filter(Method %in% c("CFU"))  %>%  mutate(Treatment = as.factor(Treatment))

# Fit the linear model
result <- aov(Survival ~ Treatment, data = filtered_data)


# Perform ANOVA
anova_result <- summary(result)

# Perform Tukey's HSD test
tukey_result <- TukeyHSD(result)

# Print the ANOVA result
print(anova_result)

# Print the Tukey's HSD test result
print(tukey_result)

```
```{r}
# Filter the data
filtered_data <- combined_data1 %>% dplyr::filter(Method %in% c("FungaLight"))  %>%  mutate(Treatment = as.factor(Treatment))

# Fit the linear model
result <- aov(Survival ~ Treatment, data = filtered_data)


# Perform ANOVA
anova_result <- summary(result)

# Perform Tukey's HSD test
tukey_result <- TukeyHSD(result)

# Print the ANOVA result
print(anova_result)

# Print the Tukey's HSD test result
print(tukey_result)

```


```{r}
# Filter the data
filtered_data <- combined_data1 %>% dplyr::filter(Method %in% c("PI"))  %>%  mutate(Treatment = as.factor(Treatment))

# Fit the linear model
result <- aov(Survival ~ Treatment, data = filtered_data)


# Perform ANOVA
anova_result <- summary(result)

# Perform Tukey's HSD test
tukey_result <- TukeyHSD(result)

# Print the ANOVA result
print(anova_result)

# Print the Tukey's HSD test result
print(tukey_result)

```