---
title: "Apply FungaLight to AmpB-treated C. glabrata cells"
author: Hanxi Tang, Bin He
date: "2025-03-31 (updated `r Sys.time()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---


```{r setup, message=FALSE}
require(flowCore)
require(flowClust)
require(openCyto)
require(tidyverse)
require(ggcyto)
require(cowplot)
#require(ggrdiges)
```

```{r}
old <- theme_set(theme_minimal(base_size = 14) + panel_border(color = "gray20"))
```

# Goal
The goal is to test FungaLight's applicability to antifungal-treated _C. glabrata_ cells.

# Data
This is the flow cytometry data for post  stained with PI+SYTO9, and run through flow cytometry (details in ELN), and plated for CFU. 

## Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "../input/20240325-29-Cgla-AmpB/"

dat0 <- read.flowSet(path = data.path, pattern = "*.fcs",
                     # the original values are already linearized. 
                     transformation = FALSE,  
                     # change parameter names to R format
                     emptyValue = FALSE,  alter.names = TRUE,
                     # only load the height variables for the fluorescent parameters
                     column.pattern = ".H|FSC|SSC")
```

Simplify the sample names

```{r}
#require(PTXQC)
oriNames <- sampleNames(dat0)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6)]
colnames(tmp) <- c("Date", "Treatment") 
sample <- data.frame(tmp) %>% 
  mutate(Treatment = 
           fct_inseq(gsub("\\.fcs", "", Treatment))
  )
rownames(sample) <- oriNames
dat = dat0 # make a copy
pData(dat) <- sample
write_tsv(pData(dat), file = "../input/20240325-Cg-AmpB-FGL-sample.tsv")
```

# Gatting strategies {.tabset}
## Load all gates from specification file
```{r}
source("../../../00-Shared/01-script/20250206-FGL-flow-gate-specification-Cg.R")
```

## Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier

```{r}
ggcyto(dat[1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```

Add gate to GS
```{r}
gs <- GatingSet(dat) # create a GatingSet
# rename the samples
sampleNames(gs) <- pData(gs)$name
# repair the Date column in pData
pData(gs)$Date = paste0("0", pData(gs)$Date)
# add the outlier gate
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

> Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat[[1]], outlier.gate)
#polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
#                    60, 105, 135,60), ncol = 2)
#colnames(polygon) <- c("FSC.H", "FSC.W")
#singlet.gate <- polygonGate(filterId = "singlet", .gate = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```

Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

## Gate for live cells
This is the live population with red fluorescence below 10^2.2

> **Main takeaway**
>
> 1. From 10mM to 1M treated samples, %live decreases and the change is much more obvious in the dual stain (PI first or simultaneous) than PI alone
>
> 2. The mock treated sample has a visible high red high green (not as high and prominent as the 100 mM treated) than 10 mM treated. It doesn't follow the trend above.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5.5), y = c(0,10^5.5)))
#polygon <- matrix(c(0, 5*10^3, 5*10^3, 0, # BL1.H, green
#                    10^2.2, 10^2.2, 0, 0),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#live.gate <- polygonGate(filterId = "live", .gate = polygon)
p.axis <- list(
  scale_x_logicle(breaks = 10^c(2,3,4,5)),
  scale_y_logicle(breaks = 10^c(2,3,4,5)),
  theme(
    strip.text = element_text(size = rel(1.1), face = 2)
  )
)
for(date in unique(sample$Date)){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(live.gate.sc) + 
    geom_stats(location = "gate", adjust = c(0.02, 2), digits = 1, ) + 
    labs(title = paste0("Date: ", date)) +
    facet_wrap(~ Treatment, ncol = 4) +# scPars +
    p.axis
  print(p)
}
```

> Gate for Red Negative Population
This is the live population with red fluorescence below 10^2.2

```{r}
```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(0, 10^3, 10^3, 0,
                    10^2.2, 10^2.2, 0, 0), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
live.gate <- polygonGate(filterId = "live", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_wrap(Treatment ~ Date, ncol = 7, labeller = labeller(.cols = label_both)) 
```
```


Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```

> Gate for Red Positive Intermediates
This is the intermediate population on the right side

```{r}
```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(10^3, 10^5, 10^5, 10^3, 
                      0, 0, 10^4, 10^2), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
inter.gate <- polygonGate(filterId = "inter", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(inter.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_wrap(Treatment ~ Date, ncol = 7, labeller = labeller(.cols = label_both)) 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, inter.gate, parent = "singlet", name = "inter")
recompute(gs)
```
```

> Gate for Red Positive Dead
This is the dead population on the right side

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(10^3, 10^2.5,10^3, 10^5,
                    10^4.5, 10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_wrap(Treatment ~ Date, ncol = 7, labeller = labeller(.cols = label_both)) 
```
```

Add this gate to the gatingSet

```{r}
```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```
```


>Exported gated event counts

```{r}
```{r}
gated_stats_f1g <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f1g$percent_Dead <- gated_stats_f1g$n_dead / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_Live <- gated_stats_f1g$n_live / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_inter <- gated_stats_f1g$n_inter / gated_stats_f1g$n_singlet
gated_stats_f1g$FGS <- 1*gated_stats_f1g$percent_Live + 0.5*gated_stats_f1g$percent_inter - gated_stats_f1g$percent_Dead
gated_data <- cbind(pData(dat.f1g), gated_stats_f1g) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
write_tsv(gated_data, file = "../Input/20240401-Cg-AmpB-gated-ht.tsv")
print(gated_data)
```
```



# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.
```{r}

```{r}
gated_data %>%
ggplot(aes(x = Treatment, y = FGS)) + 
  geom_boxplot()+
  geom_point()+
  theme_minimal() +
  labs(x = "Treatment AmphB (ng/ml)",
       y = "Gated Live Population Percentage") 

#ggsave("../output/20240117-dye-Order-Flow-Plot.png", width = 7, height = 5)
```
```

```{r}
```{r}
combined_data <- read_tsv("../Input/20240401-Cg-AmpB-gated-CFU-ht.tsv") %>% 
mutate(Treatment = factor(Treatment, levels = c(0, "5C", 0.6, 6, 1000))) %>% 

p1 <- combined_data %>% 
   dplyr::filter(!(Treatment %in% c("1", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 2, position = position_jitter(0.05)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  ggtitle("PI only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot() 
  

p2 <- combined_data %>% 
   dplyr::filter(!(Treatment %in% c("1", "0")))  %>%
  ggplot(aes(x = Treatment, y = CFU_Survival))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 2, position = position_jitter(0.05)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  ggtitle("CFU") +  
  theme(plot.title = element_text(hjust = 15)) +
  theme_cowplot()

combined_plot <- plot_grid(p1, p2, nrow = 1, ncol = 3, common.legend = TRUE, align = 'hv')

print(combined_plot)
#ggsave("../output/20240329-Ca-Log-Gated.png")
```
Correlation Graph Test
```{r}
combined_data %>%
  dplyr::filter(!(Treatment %in% c("1", "0")))   %>% 
  ggplot(aes(x = percent_Live, y = CFU_Survival, color = Treatment)) + geom_point() + stat_smooth(method = "lm")
ggplot(combined_data, aes(x = percent_Dead, y = CFU_Survival)) +  geom_point ()
ggplot(combined_data, aes(x = percent_inter, y = CFU_Survival)) +  geom_point ()
ggplot(combined_data, aes(x = FGS, y = CFU_Survival)) +  geom_point ()
```

> Output graph

```{r}
combined_data %>%
  mutate(Treatment = paste(Treatment, "ng/ml", sep = " ")) %>% 
  dplyr::filter(!(Treatment %in% c("1 ng/ml", "0 ng/ml")))   %>% 
ggplot(aes(x = percent_Live, y = CFU_Survival)) + 
  geom_point(aes(color = Treatment), size = 3) +
  stat_smooth(method = "lm") +
  labs(x = "Gated Live Population Percentage",
       y = "CFU Survival")  +
  theme_minimal() +
  theme(axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(color = "black"),
        legend.text = element_text(size = 14, face = "bold"), 
        legend.title = element_text(size = 14, face = "bold")) + scale_color_viridis(discrete = TRUE)

ggsave("../output/20240402-Cg-AmphB.png")
```






