---
title: "Apply FungaLight to AmpB-treated C. glabrata cells"
author: Hanxi Tang, Bin He
date: "2025-03-31 (updated `r Sys.time()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---


```{r setup, message=FALSE}
require(flowCore)
require(flowClust)
require(openCyto)
require(tidyverse)
require(ggcyto)
require(cowplot)
#require(ggrdiges)
```

```{r}
old <- theme_set(theme_minimal(base_size = 14) + panel_border(color = "gray20"))
```

# Goal
The goal is to test FungaLight's applicability to antifungal-treated _C. glabrata_ cells.

# Data
This is the flow cytometry data for post  stained with PI+SYTO9, and run through flow cytometry (details in ELN), and plated for CFU. 

## Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "../input/20240325-29-Cgla-AmpB/"

dat0 <- read.flowSet(path = data.path, pattern = "*.fcs",
                     # the original values are already linearized. 
                     transformation = FALSE,  
                     # change parameter names to R format
                     emptyValue = FALSE,  alter.names = TRUE,
                     # only load the height variables for the fluorescent parameters
                     column.pattern = ".H|FSC|SSC")
```

Simplify the sample names

```{r}
#require(PTXQC)
oriNames <- sampleNames(dat0)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6)]
colnames(tmp) <- c("Date", "Treatment") 
sample <- data.frame(tmp) %>% 
  mutate(
    Treatment = fct_inseq(gsub("\\.fcs", "", Treatment)),
    name = paste(Date, Treatment, sep = "_")
  )
rownames(sample) <- oriNames
dat <- dat0 # make a copy
pData(dat) <- sample
write_tsv(pData(dat), file = "../input/20240325-Cg-AmpB-FGL-sample.tsv")
```

# Gatting strategies {.tabset}
## Load all gates from specification file
```{r}
source("../../../00-Shared/01-script/20250206-FGL-flow-gate-specification-Cg.R")
```

## Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier

```{r}
ggcyto(dat[1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```

Add gate to GS
```{r}
gs <- GatingSet(dat) # create a GatingSet
# rename the samples
sampleNames(gs) <- pData(gs)$name
# repair the Date column in pData
pData(gs)$Date = paste0("0", pData(gs)$Date)
# add the outlier gate
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

> Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat[[1]], outlier.gate)
#polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
#                    60, 105, 135,60), ncol = 2)
#colnames(polygon) <- c("FSC.H", "FSC.W")
#singlet.gate <- polygonGate(filterId = "singlet", .gate = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```

Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

## Gate for live cells
This is the live population with red fluorescence below 10^2.2

> **Main takeaway**
>
> 1. From 10mM to 1M treated samples, %live decreases and the change is much more obvious in the dual stain (PI first or simultaneous) than PI alone
>
> 2. The mock treated sample has a visible high red high green (not as high and prominent as the 100 mM treated) than 10 mM treated. It doesn't follow the trend above.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5.5), y = c(0,10^5.5)))
#polygon <- matrix(c(0, 5*10^3, 5*10^3, 0, # BL1.H, green
#                    10^2.2, 10^2.2, 0, 0),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#live.gate <- polygonGate(filterId = "live", .gate = polygon)
p.axis <- list(
  scale_x_logicle(breaks = 10^c(2,3,4,5)),
  scale_y_logicle(breaks = 10^c(2,3,4,5)),
  theme(
    strip.text = element_text(size = rel(1.1), face = 2)
  )
)
for(date in unique(sample$Date)){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(live.gate.AmpB) + 
    geom_stats(location = "gate", adjust = c(0.05, 2), digits = 1, ) + 
    labs(title = paste0("Date: ", date)) +
    facet_wrap(~ Treatment, ncol = 4) +# scPars +
    p.axis
  print(p)
}
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate.AmpB, parent = "singlet", name = "live")
recompute(gs)
```

## Gate for the intermediate population with high SYTO9
This is the intermediate population with high Red and Green. We hypothesize that they represent oxidatively damaged cells with partially compromised plasma membranes, resulting in more SYTO9 accumulation but not a significant increase in PI.

```{r}
for(date in unique(sample$Date)){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H),
              subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(inter.gate.AmpB) + 
    geom_stats(location = "data",  adjust = c(0.05, 5), digits = 3) + 
    labs(title = paste0("Date: ", date)) +
    facet_wrap(~ fct_inseq(Treatment), nrow = 2) + 
    scPars + p.axis
  print(p)
}
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, inter.gate.AmpB, parent = "singlet", name = "intermediate")
recompute(gs)
```

## Gate for dead cells

```{r}
for(date in unique(sample$Date)){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H),
              subset = "singlet") + 
    geom_hex(bins = 128) + 
    #geom_gate(live.gate.AmpB) + 
    #geom_gate(inter.gate.AmpB) + 
    geom_gate(dead.gate.AmpB) + 
    geom_stats(location = "gate",  adjust = c(0.05, 1), digits = 3) + 
    labs(title = paste0("Date: ", date)) +
    facet_wrap(~ fct_inseq(Treatment), nrow = 2) + 
    scPars + p.axis
  print(p)
}
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate.AmpB, parent = "singlet", name = "dead")
recompute(gs)
```

## Representative FGL staining pattern
Since we modified the gates, will need to show the gating strategy again for AmpB treated _C. glabrata_.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5.5), y = c(0,10^4.5)))

p <- ggcyto(gs[with(pData(gs), Date == "032924" & Treatment != "0.02")], 
            aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
  geom_hex(bins = 128) + 
  geom_gate(live.gate.AmpB, colour = "gray20") + 
  geom_gate(inter.gate.AmpB, colour = "steelblue3") + 
  #geom_gate(dead.gate.AmpB, colour = "orange2") + 
  #geom_stats(location = "gate", adjust = c(0.02, 2), digits = 1, fill = NA) + 
  facet_wrap(~ Treatment, nrow = 1,
             labeller = labeller(Treatment = \(x) paste0(x, " ng/mL"))) + 
  scPars + p.axis +
  theme(plot.title = element_blank(), axis.title = element_blank(),
        strip.text = element_text(size = rel(0.8)))
p

ggsave("../output/20250401-Cg-AmpB-FGL-representative-plot.png", 
       height = 2, width = 8)
```

## Explore cell size difference in hydrogen peroxide treated cells
This is based on Jinye's suggestion:
> I am wondering is whether the cell size and granularity would affect the staining in different ways among species. Thus, I think it makes sense to plot the FL on FSC or SSC and perform a linear regression analysis for each species, if there is no significant difference, we can say the cell property of different species does not affect the staining in different ways. This is a simplified or adapted idea from PMID: 21952134.

Let's plot the distribution of FSC.H as a proxy for cell size
```{r}
mult_format <- function() {
     function(x) format(x/10000,digits = 2) 
}
```

```{r}
p1 <- ggcyto(gs, aes(x = FSC.H), subset = "singlet") + 
  ggridges::geom_density_ridges(
    aes(y = Treatment),
        fill = "forestgreen", alpha = 0.8) + 
  scale_x_continuous(labels = mult_format(), name = "FSC.H x 10000") +
  facet_wrap(~.) + labs(title = NULL, y = "Amphotericin B (ng/mL)") +
  theme(strip.text = element_blank())
p1
ggsave("../output/20250401-Cg-AmpB-cell-size-distribution-compare.png",
       width = 3, height = 4)
```

> There is a leftward shift but no reverse shift at the lowest dose.

# Extract gated stats and MFI

% of events in each of the three gates
```{r}
# we are only interested in the populations after singlet gating
nodes <- c("singlet", "live", "intermediate", "dead")
# get the event count
gated_cnt <- gs_pop_get_stats(gs, nodes, type = "count")
# get the MFI for each parameter in each population of interest
gated_mfi <- gs_pop_get_stats(gs, nodes, type = pop.MFI) %>% 
  select(sample, pop, starts_with("FungaLight"))
```

Combine the data and meta data
```{r}
gated_stats <- full_join(gated_cnt, gated_mfi, by = c("sample", "pop")) %>% 
  right_join(pData(gs), by = c("sample" = "name")) %>% 
  relocate(Date, Treatment, .after = sample) %>% 
  relocate(sample, .after = last_col())
write_tsv(gated_stats, file = "../output/20250401-Cg-AmpB-FGL-gated-stats.tsv")

# we don't need the MFI for the latter analysis. add meta data to gated_cnt
gated_freq <- left_join(sample, gated_cnt, by = c("name" = "sample")) %>% 
  select(-name) %>% 
  group_by(Date, Treatment) %>% 
  mutate(perc = count / count[pop == "singlet"], .after = count)
```

# Compare with CFU
## Read CFU data 
These are averaged CFU provided by Hanxi. Here are the raw cfu data
```{r}
cfu_raw <- read_tsv("../input/20240325-29-Cgla-AmpB-CFU-raw.tsv",
                    col_types = "cccccciiii", comment = "#")
cfu_calc <- cfu_raw %>% 
  select(Date, Species, Treatment = AmpB_ng_mL, Group, 
         Dilution, Raw_count = Total) %>% 
  mutate(Date = ymd(Date) |> format("%m%d%y"),
         Treatment = fct_inseq(Treatment),
         CFU = Dilution * Raw_count) %>% 
  # group by Date to calculate the ratio of MO to MM
  group_by(Date) %>% 
  mutate(p_live = CFU / CFU[Group == "MM"])

# export the data for paper
write_tsv(cfu_calc, file = "../output/20250325-Cg-AmpB-cfu-estimate.tsv")
```

## Calculate survival from FungaLight

For FungaLight, we will calculate the percent live, intermediate and dead events.
```{r}
dye_based <- gated_freq %>% 
  # record the total number of events as the count of singlets
  group_by(Date, Treatment) %>% 
  mutate(total_event_count = count[pop == "singlet"]) %>% 
  # now we don't need the singlet population and the count variable
  select(-count) %>% 
  dplyr::filter(pop != "singlet") %>%
  #ungroup() %>% 
  # we will pivot this data frame to a wider format, making 
  # two new variables out of the percentage variable
  pivot_wider(names_from = pop, values_from = perc, names_prefix = "p_") %>% 
  #mutate(p_sum = p_live + p_intermediate + p_dead) %>% 
  relocate(p_live, .after = total_event_count)

# export the data for paper
write_tsv(dye_based, file = "../output/20250325-Cg-AmpB-FGL-estimate.tsv")
```

Combine the dye based results with cfu
```{r}
cfu_dye <- bind_rows(
  CFU = select(cfu_calc, Date, Treatment, p_live),
  FungaLight = select(dye_based, Date, Treatment, starts_with("p")),
  .id = "Method"
)
```

## Plotting the results
### CFU survival
```{r}
p.cfu <- cfu_calc %>% 
  dplyr::filter(Treatment != "0.02") %>%
  ggplot(aes(x = Treatment, y = p_live)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray70", width = 0.8) +
  geom_point(aes(shape = as.character(Date)), 
             position = position_jitter(0.1),
             size = 1.5) + 
  scale_shape_manual(values = 4:7, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.05), breaks = seq(0,1,0.5)) +
  labs(title = "CFU", x = "AmpB (ng/mL)", y = NULL) +
  theme(
    plot.title = element_text(size = rel(0.8), hjust = 0.5,
                              margin = margin(b = 2)),
    axis.text = element_text(size = rel(0.7)),
    axis.title = element_text(size = rel(0.7))
  )

p.cfu
ggsave("../output/20250401-Cg-AmpB-CFU.png", width = 3.5, height = 1.8)
```
### FungaLight
```{r}
p.dye.side_by_side <- dye_based %>% 
  dplyr::filter(Treatment != "0.02") %>%
  select(Date, Treatment, p_live, p_intermediate) %>% 
  pivot_longer(cols = c(p_live, p_intermediate), 
               names_to = "par", values_to = "perc") %>% 
  mutate(par = factor(par, levels = c("p_live", "p_intermediate"),
                      labels = c("% live", "% intermediate"))) %>% 
  ggplot(aes(x = fct_inseq(Treatment), y = perc, group = par)) +
  geom_bar(aes(fill = par), position = position_dodge(0.9),
           stat = "summary", fun = "mean", width = 0.8) +
  geom_point(aes(shape = as.character(Date), color = par), 
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.9),
             size = 1.2) + 
  scale_fill_manual(NULL, values = c("gray70", "lightblue2")) +
  scale_color_manual(NULL, values = c("black", "steelblue")) +
  scale_shape_manual(values = 4:7, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.05), breaks = seq(0,1,0.5)) +
  labs(title = "FungaLight", x = "AmpB (ng/mL)", y = NULL) +
  theme(
    plot.title = element_text(size = rel(0.8), hjust = 0.5,
                              margin = margin(b = 2)),
    axis.text = element_text(size = rel(0.7)),
    axis.title = element_text(size = rel(0.7)),
    legend.position = "none"
  )

p.dye.side_by_side
ggsave("../output/20250401-Cg-AmpB-FGL.png", width = 3.5, height = 1.8)
```
Assemble the plot
```{r}
plot_grid(
  p.cfu + labs(x = NULL, y = NULL) +
    theme(plot.title = element_text(size = rel(0.8),
                                    face = 2, margin = margin(b = 2)),
          axis.text = element_text(size = rel(0.7)),
          axis.text.x = element_blank()),
  NULL,
  p.dye.side_by_side + labs(x = NULL, y = NULL) +
    theme(#axis.text.y = element_blank(),
          plot.title = element_text(size = rel(0.8),
                                    face = 2, margin = margin(b = 1)),
          axis.text = element_text(size = rel(0.7)),
          legend.position = "none"),
  nrow = 3, rel_heights = c(1, 0.3, 1.12)
)
ggsave("../output/20250401-Cg-AmpB-CFU-FGL-compare.png",
       width = 3.5, height = 3.5)
```

### Compare CFU vs FGL
We will focus on comparing the two methods for their p_live estimates
```{r}
tmp <- cfu_dye %>% 
  select(-p_intermediate, -p_dead) %>% # remove unwanted variables
  pivot_wider(names_from = Method, values_from = p_live) %>% 
  dplyr::filter(!Treatment %in% c("0", "0.02")) # remove mock and 0.02 ng/mL
# removing mock because CFU for mock is used to normalize the treatment groups

p <- ggplot(tmp, aes(x = CFU, y = FungaLight)) + 
  geom_abline(slope = 1, intercept = 0, linetype = 2,
              color = "gray20", linewidth = 0.5) +
  geom_point(aes(color = Treatment), size = 2) +
  scale_color_brewer("AmpB (ng/mL)", palette = "Set1") +
  scale_x_continuous(labels = scales::percent, 
                     limits = c(NA, 1.05), breaks = seq(0,1,0.25)) +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.05), breaks = seq(0,1,0.25)) +
  theme(
    legend.title = element_text(size = rel(0.8)),
    legend.text = element_text(size = rel(0.8))
  )


p
ggsave("../output/20240402-Cg-AmpB-CFU-vs-FGL-p_live-scatter.png",
       width = 4.2, height = 2.5)
```

## Statistical tests
Calculate Pearson's and Spearman's correlation between the p_live estimates using CFU vs FGL
```{r}
tmp <- cfu_dye %>% 
  select(-p_intermediate, -p_dead) %>% # remove unwanted variables
  pivot_wider(names_from = Method, values_from = p_live) %>% 
  dplyr::filter(!Treatment %in% c("0", "0.02")) # remove mock and 0.02 ng/mL

corr_p_live = c(
  Pearson = cor(tmp$CFU, tmp$FungaLight, method = "pearson"),
  Spearmann = cor(tmp$CFU, tmp$FungaLight, method = "spearman")
)

print(corr_p_live)
```

```{r}
tmp <- aov(CFU ~ Treatment, data = dplyr::filter(cfu_calc, Treatment != "0.02"))
TukeyHSD(tmp)
```

For p_live with FungaLight
```{r}
tmp <- dye_based %>% 
  dplyr::filter(Treatment != "0.02") %>%
  aov(p_live ~ fct_inseq(Treatment), data = .)
TukeyHSD(tmp)
```

For p_intermediate with FungaLight
```{r}
tmp <- dye_based %>% 
  dplyr::filter(Treatment != "0.02") %>%
  aov(p_intermediate ~ fct_inseq(Treatment), data = .)
TukeyHSD(tmp)
```
For p_dead with FungaLight
```{r}
tmp <- dye_based %>% 
  dplyr::filter(Treatment != "0.02") %>%
  aov(p_dead ~ fct_inseq(Treatment), data = .)
TukeyHSD(tmp)
```

# Output Graph Production

```{r}
gated_data %>%
ggplot(aes(x = Treatment, y = FGS)) + 
  geom_boxplot()+
  geom_point()+
  theme_minimal() +
  labs(x = "Treatment AmphB (ng/ml)",
       y = "Gated Live Population Percentage") 

#ggsave("../output/20240117-dye-Order-Flow-Plot.png", width = 7, height = 5)
```
```

```{r}
```{r}
combined_data <- read_tsv("../Input/20240401-Cg-AmpB-gated-CFU-ht.tsv") %>% 
mutate(Treatment = factor(Treatment, levels = c(0, "5C", 0.6, 6, 1000))) %>% 

p1 <- combined_data %>% 
   dplyr::filter(!(Treatment %in% c("1", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 2, position = position_jitter(0.05)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  ggtitle("PI only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot() 
  

p2 <- combined_data %>% 
   dplyr::filter(!(Treatment %in% c("1", "0")))  %>%
  ggplot(aes(x = Treatment, y = CFU_Survival))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightgreen", alpha = 0.7) +
  geom_point(size = 2, position = position_jitter(0.05)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  ggtitle("CFU") +  
  theme(plot.title = element_text(hjust = 15)) +
  theme_cowplot()

combined_plot <- plot_grid(p1, p2, nrow = 1, ncol = 3, common.legend = TRUE, align = 'hv')

print(combined_plot)
#ggsave("../output/20240329-Ca-Log-Gated.png")
```
Correlation Graph Test
```{r}
combined_data %>%
  dplyr::filter(!(Treatment %in% c("1", "0")))   %>% 
  ggplot(aes(x = percent_Live, y = CFU_Survival, color = Treatment)) + geom_point() + stat_smooth(method = "lm")
ggplot(combined_data, aes(x = percent_Dead, y = CFU_Survival)) +  geom_point ()
ggplot(combined_data, aes(x = percent_inter, y = CFU_Survival)) +  geom_point ()
ggplot(combined_data, aes(x = FGS, y = CFU_Survival)) +  geom_point ()
```

> Output graph

```{r}
combined_data %>%
  mutate(Treatment = paste(Treatment, "ng/ml", sep = " ")) %>% 
  dplyr::filter(!(Treatment %in% c("1 ng/ml", "0 ng/ml")))   %>% 
ggplot(aes(x = percent_Live, y = CFU_Survival)) + 
  geom_point(aes(color = Treatment), size = 3) +
  stat_smooth(method = "lm") +
  labs(x = "Gated Live Population Percentage",
       y = "CFU Survival")  +
  theme_minimal() +
  theme(axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(color = "black"),
        legend.text = element_text(size = 14, face = "bold"), 
        legend.title = element_text(size = 14, face = "bold")) + scale_color_viridis(discrete = TRUE)

ggsave("../output/20240402-Cg-AmphB.png")
```






