---
title: "Compare CFU and FungaLight for log dilution of H2O2, cell-size normalized"
author: Bin He
date: "2024-04-09 (updated `r Sys.time()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
#require(ggrdiges)
```

```{r}
old <- theme_set(theme_minimal(base_size = 14) + panel_border(color = "gray20"))
```

# Background & Goal
This is a duplicate of the similarly-named R markdown on 2024-03-21. The goal of this modified analysis is to test the effect of normalizing the fluorescence levels by cell size.

## original background and gaol
Having optimized the FungaLight assay conditions, we would like to compare its estimates to those by CFU and PI alone, two commonly used methods in the literature.

The experiments for _C. glabrata_ were conducted on 2024.02.14-17. Post-treatment cells were stained with PI or PI+SYTO9, and run through flow cytometry. The same sample was also plated for CFU.

A similar experiment was done for _S. cerevisiae_ on 2024.02.25-28.

# Import data
> FCS files are stored in rdss/Users/Hanxi/flow cytometry under the above dates. They are copied to this repository under the same names.

_note_: samples are labeled by the H2O2 doses they were treated with. an exception is "5C", which refers to samples taken prior to the treatment and stored in the fridge at 5 Celsius until treated samples are ready for flow cytometer. this was done as a control experiment. ask Hanxi about the specific rationale. 

```{r}
# use relative path to make it easier for collaboration
data.path = "../input/20240214-17 Log H2O2 correlation Cg/"
dat0 <- read.flowSet(path = data.path, pattern = "*.fcs",
                     transformation = FALSE,  # the original values are already linearized. 
                     emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                     column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
oriNames <- sampleNames(dat0)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6, 7)]
colnames(tmp) <- c("Date", "Treatment", "Dye")
#treatment.levels <- c("mock", "5C", "10", "100", "1000")
sample <- data.frame(tmp) %>% 
  mutate(
    Dye = ifelse(Dye == "p.fcs", "PI", ifelse(Dye == "b.fcs", "Both", Dye)), 
    Treatment = fct_inorder(Treatment),
    name = paste(Date, paste0(Treatment, 
                              ifelse(Treatment == "5C", "", "mM")), 
                 Dye, sep = "_")
  )
rownames(sample) <- oriNames
dat <- dat0 # make a copy
pData(dat) <- sample
write_tsv(pData(dat), file = "../input/20240214-log-h2o2-cg-sample.tsv")
```


# Calculate cell normalized fluorescence

Following the example in the [flowCore manual](https://bioconductor.org/packages/release/bioc/vignettes/flowCore/inst/doc/HowTo-flowCore.pdf), we first create a transformation function, then set up a `transformList` object, and finally use it to transform the flowSet _before_ constructing the gatingSet
```{r}
mfsc <- 5e5 # this is just used to make the new variable on a similar scale as 
            # the original one, by making the denominator ~1
# Create new parameters: BL1.H / FSC.H^1.5 and BL3.H / FSC.H^1.5
# this is not recommended for programmatic use based on flowCore manual,
# but the recommended transformList approach requires writing hard-to-read
# custom functions. let's see if this would work.

# FSC.H^{3/2} is roughly the volume of a sphere. Here we use an exponent of
# 1.2 for BL1.H and 1 for BL3.H based on empirical information (see below)
ndat <- transform(dat,
                  BL1.H = `BL1.H` / ((`FSC.H`/mfsc)^1.2),
                  BL3.H = `BL3.H` / ((`FSC.H`/mfsc)^0.8))
```

For the BL1.H channel, we can see below that the 'cell size normalized' fluorescence values are less correlated with FSC when the sample contains highly fluorescent cells (damaged or dead).
```{r}
p1 <- ggcyto(dat[pData(dat)$Treatment == "100" & pData(dat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL1.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "BL1.H")

p2 <- ggcyto(ndat[pData(ndat)$Treatment == "100" & pData(ndat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL1.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "normalized BL1")
plot_grid(as.ggplot(p1), as.ggplot(p2), nrow = 2)
```
And also in mock treated sample.
In samples with lowly fluorescent populations, the correction actually induces a negative correlation between the fluorescence signals and FSC, even if I set the exponent to 1.
```{r}
p1 <- ggcyto(dat[pData(dat)$Treatment == "0" & pData(dat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL1.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "BL1.H")

p2 <- ggcyto(ndat[pData(ndat)$Treatment == "0" & pData(ndat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL1.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "normalized BL1")
plot_grid(as.ggplot(p1), as.ggplot(p2), nrow = 2)
```
For the BL3.H channel, we can see below that the 'cell size normalized' fluorescence values are less correlated with FSC when the sample contains highly fluorescent cells (damaged or dead).
```{r}
p1 <- ggcyto(dat[pData(dat)$Treatment == "100" & pData(dat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL3.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "BL3.H")

p2 <- ggcyto(ndat[pData(ndat)$Treatment == "100" & pData(ndat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL3.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "normalized BL3")
plot_grid(as.ggplot(p1), as.ggplot(p2), nrow = 2)
```

In samples with lowly fluorescent populations, the correction could induce a negative correlation between the fluorescence signals and FSC, even if I set the exponent to 1.
```{r}
p1 <- ggcyto(dat[pData(dat)$Treatment == "0" & pData(dat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL3.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "BL3.H")

p2 <- ggcyto(ndat[pData(ndat)$Treatment == "0" & pData(ndat)$Dye == "Both"], 
             aes(x = FSC.H, y = BL3.H)) + 
  geom_hex(bins = 64) + scale_y_logicle() +
  scale_x_continuous(labels = scales::label_number(scale = 1e-4)) +
  stat_smooth(method = "lm", formulat = y~x) +
  labs(x = "FSC.H", y = "normalized BL3")
plot_grid(as.ggplot(p1), as.ggplot(p2), nrow = 2)
```

# Gatting strategies {.tabset}
## Load all gates from specification file
```{r}
source("../script/20250206-FGL-flow-gate-specification.R")
```

## Gating for singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.
>Gate for outlier 

```{r}
#outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(ndat[1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(ndat) # create a GatingSet
# rename the samples
sampleNames(gs) <- pData(gs)$name
# repair the Date column in pData
pData(gs)$Date = paste0("0", pData(gs)$Date)
# add the outlier gate
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

> Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(ndat[[1]], outlier.gate)
#polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
#                    60, 105, 135,60), ncol = 2)
#colnames(polygon) <- c("FSC.H", "FSC.W")
#singlet.gate <- polygonGate(filterId = "singlet", .gate = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

## Gate for live cells
This is the live population with red fluorescence below 10^2.2

> **Main takeaway**
>
> 1. From 10mM to 1M treated samples, %live decreases and the change is much more obvious in the dual stain (PI first or simultaneous) than PI alone
>
> 2. The mock treated sample has a visible high red high green (not as high and prominent as the 100 mM treated) than 10 mM treated. It doesn't follow the trend above.

```{r}
scPars <- ggcyto_par_set(limits = list(x = c(0,10^5.5), y = c(0,10^5.5)))
#polygon <- matrix(c(0, 5*10^3, 5*10^3, 0, # BL1.H, green
#                    10^2.2, 10^2.2, 0, 0),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#live.gate <- polygonGate(filterId = "live", .gate = polygon)
p.axis <- list(
  scale_x_logicle(breaks = 10^c(2,3,4,5)),
  scale_y_logicle(breaks = 10^c(2,3,4,5)),
  theme(
    strip.text = element_text(size = rel(1.1), face = 2)
  )
)
for(date in c("021424", "021624", "021724")){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H), subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(live.gate) + 
    geom_stats(location = "data", adjust = c(0.005, 5), digits = 1) + 
    labs(title = paste0("Date: ", date)) +
    facet_grid(Dye ~ Treatment) +# scPars +
    p.axis
  print(p)
}
```

Add this gate to the gatingSet

```{r}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```
## Gate for the intermediate population with high SYTO9
This is the intermediate population with high Red and Green. We hypothesize that they represent oxidatively damaged cells with partially compromised plasma membranes, resulting in more SYTO9 accumulation but not a significant increase in PI.

> <span style="color:red">Important: I changed the bottom boundary from 10^2.2 to 10^2.3, and the left boundary from 10^3 to 2*10^3</span>

> **Main takeaway**
>
> 1. Different from the %live gate, the intermediate population is most obvious in the 
> 100 mM treated sample, and absent in both the 10 mM and 1M treated ones. This suggests
> that this population could be useful for distinguishing stress levels below the lethal concentration
> but above the tolerance levels of the species - in this case, < 1M and > 10 mM.
>
> 2. The mock treated sample showed an intermediate population. Was unexpected.

```{r}
#polygon <- matrix(c(10^5, 1*10^3, 10^5,
#                    10^2.2, 10^2.2, 10^4), ncol = 2)
#polygon <- matrix(c(10^3, 10^5.5, 10^5.5,   # BL1.H, green
#                    10^2.2, 10^2.2, 10^4.5),# BL3.H, red
#                  ncol = 2)
#colnames(polygon) <- c("BL1.H", "BL3.H")
#inter.gate <- polygonGate(filterId = "inter", .gate = polygon)
for(date in c("021424", "021624", "021724")){
  p <- ggcyto(gs[pData(gs)$Date == date], aes(x = BL1.H, y = BL3.H),
              subset = "singlet") + 
    geom_hex(bins = 128) + 
    geom_gate(inter.gate) + 
    geom_stats(location = "data",  adjust = c(0.05, 5), digits = 3) + 
    labs(title = paste0("Date: ", date)) +
    facet_grid(Dye ~ Treatment) + scPars + 
    p.axis
  print(p)
}
```

```{r}
gs_pop_add(gs, inter.gate, parent = "singlet", name = "intermediate")
recompute(gs)
```

# Extract gated stats and MFI

% of events in each of the three gates
```{r}
# we are only interested in the populations after singlet gating
nodes <- c("singlet", "intermediate", "live")
# get the event count
gated_cnt <- gs_pop_get_stats(gs, nodes, type = "count")
# get the MFI for each parameter in each population of interest
gated_mfi <- gs_pop_get_stats(gs, nodes, type = pop.MFI) %>% 
  select(sample, pop, starts_with("FungaLight"))
```

Combine the data and meta data
```{r}
gated_stats <- full_join(gated_cnt, gated_mfi, by = c("sample", "pop")) %>% 
  right_join(pData(gs), by = c("sample" = "name")) %>% 
  relocate(Date, Treatment, Dye, .after = sample) %>% 
  relocate(sample, .after = last_col())
write_tsv(gated_stats, file =
            "../output/20250410-log-h2o2-Cg-gated-stats-cell-size-normalized.tsv")

# we don't need the MFI for the latter analysis. add meta data to gated_cnt
gated_freq <- left_join(sample, gated_cnt, by = c("name" = "sample")) %>% 
  select(-name) %>% 
  group_by(Date, Treatment, Dye) %>% 
  mutate(perc = count / count[pop == "singlet"], .after = count)
```

# Compare with CFU
## Read CFU data 
These are averaged CFU provided by Hanxi. The raw counts data will be provided by her separately.
```{r}
cfu <- read_tsv("../input/20240219-log-h2o2-Cg-CFU-HT.tsv") %>% 
  mutate(Method = "CFU", .after = Treatment) %>% 
  mutate(Treatment = fct_recode(Treatment, `0` = "mock")) %>%
  rename(p_live = CFU)
```

Here are the raw cfu data
```{r}
cfu_raw <- read_tsv("../input/20240214-0405-h2o2-Cg-CFU-raw.tsv",
                    col_types = "cccccciiii", comment = "#")
cfu_calc <- cfu_raw %>% 
  select(Date, Treatment = H2O2_mM, Group, Dilution, Raw_count = Total) %>% 
  mutate(Date = ymd(Date),
         Treatment = fct_inseq(Treatment),
         CFU = Dilution * Raw_count) %>% 
  # group by Date to calculate the ratio of MO to MM
  group_by(Date) %>% 
  mutate(p_live = CFU / CFU[Group == "MM"]) %>% 
  # we only need the log dilution series
  dplyr::filter(Date < ymd("20240301"))

# export the data for paper
#write_tsv(cfu_calc, file = "../output/20250226-log-h2o2-cfu-estimate.tsv")
```

## Calculate survival from PI alone or FungaLight

For both PI alone and FungaLight, we will calculate the percent live events. For FungaLight specifically, we will also record the percent intermediate events and will investigate its utility
```{r}
dye_based <- gated_freq %>% 
  # record the total number of events as the count of singlets
  group_by(Date, Treatment, Dye) %>% 
  mutate(total_event_count = count[pop == "singlet"]) %>% 
  # now we don't need the singlet population and the count variable
  select(-count) %>% 
  dplyr::filter(pop != "singlet") %>%
  #ungroup() %>% 
  # we will pivot this data frame to a wider format, making 
  # two new variables out of the percentage variable
  pivot_wider(names_from = pop, values_from = perc, names_prefix = "p_") %>% 
  mutate(p_dead = 1 - p_live - p_intermediate) %>% 
  relocate(p_live, .after = total_event_count)

# export the data for paper
write_tsv(dye_based, file =
            "../output/20250410-log-h2o2-fgl-cell-size-normalized-estimate.tsv")
```

Combine the dye based results with cfu
```{r}
cfu_dye <- bind_rows(cfu,
  select(dye_based, Date, Treatment, Method = Dye, starts_with("p"))) %>% 
  mutate(
    Method = factor(Method, levels = c("CFU", "PI", "Both"),
                    labels = c("CFU", "PI", "FungaLight")),
    Treatment = fct_inorder(Treatment)
  )
```

## Plotting the results
Compare the three methods
```{r}
h2o2.levels <- c("0" = "0", "10" = "10", "100" = "100", "1000" = "1K")
p <- cfu_dye %>% 
  dplyr::filter(Treatment != "5C") %>% 
  ggplot(aes(x = Treatment, y = p_live, group = Method)) +
  geom_bar(stat = "summary", fun = "mean", fill = "gray70", width = 0.7) +
  geom_point(aes(shape = Date), position = position_jitter(0.2),
             size = 1.5) + 
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  scale_x_discrete(labels = h2o2.levels) +
  #coord_flip() + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  facet_wrap(~Method, nrow = 1) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = rel(0.9), face = 1),
        axis.text.x = element_text(size = rel(0.9), angle = 90, vjust = .5),
        axis.text.y = element_text(size = rel(0.9)),
        axis.title = element_text(size = rel(0.9)),
        panel.spacing = unit(1, "lines"))
p
#ggsave("../output/20250116-perc-live-compare-three-methods.png",
#       width = 4, height = 2.5)
```
# Compare with non-normalized FGL estimates
Let's read in the previous unnormalized estimates
```{r}
dye_based_orig <- read_tsv("../output/20250226-log-h2o2-fgl-estimate.tsv")
dye_based_combine <- bind_rows(
  Normalized = dye_based,
  Non_normalized = dye_based_orig,
  .id = "Cell_size"
) %>% dplyr::filter(Treatment != "5C", Dye == "Both")
```

Compare the results
```{r}
dye_based_combine %>% 
  group_by(Cell_size, Treatment) %>% 
  summarize(across(c(p_live, p_intermediate), 
                   list(mean = mean, sd = sd)))
```

```{r}
pd <- position_jitterdodge(dodge.width = 0.9, jitter.width = 0.1)
p <- dye_based_combine %>% 
  dplyr::filter(Treatment != "5C", Dye == "Both") %>% 
  ggplot(aes(x = Treatment, y = p_live, group = Cell_size)) +
  #stat_summary(fun.data = "mean_sdl", fun.args = list(mult = 1), position = pd) +
  geom_point(aes(shape = Date, color = Cell_size), position = pd, size = 2) + 
  scale_shape_manual(values = 4:6, guide = "none") +
  scale_color_manual("Cell size", values = c("gray20", "orange")) +
  scale_y_continuous(labels = scales::percent, 
                     limits = c(NA, 1.1), breaks = seq(0,1,0.5)) +
  scale_x_discrete(labels = h2o2.levels) +
  #coord_flip() + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live") +
  #facet_wrap(~Method, nrow = 1) +
  theme(legend.title = element_text(size = rel(0.8)),
        legend.text = element_text(size = rel(0.7)))
p
ggsave("../output/20250410-perc-live-compare-cell-normalized-or-not.png",
       width = 5, height = 2.5)
```


## Statistical tests
For p_live with FungaLight before normalization
```{r}
tmp <- dye_based_orig |>
  dplyr::filter(Treatment != "5C", Dye == "Both")

test <- aov(p_live ~ Treatment, data = tmp)
TukeyHSD(test)
```

For p_live with FungaLight after normalization
```{r}
tmp <- dye_based |>
  dplyr::filter(Treatment != "5C", Dye == "Both")

test <- aov(p_live ~ Treatment, data = tmp)
TukeyHSD(test)
```